
<div>
<img src=img/logo_deployer.png alt="recon_deployer_logo" width="600px">
</div>

<br>


# Welcome to the *RECON deployer*!

The *RECON deployer* is a project of the R Epidemics Consortium
([RECON](http://www.repidemicsconsortium.org)) aiming to provide a portable,
internet-free R environment for epidemics analysis. In a *deployer* folder, you
will find the latest version of R and associated tools (RStudio, Rtools, ...),
as well as over 1,000 packages for data analysis, including all RECON packages,
some other useful packages for epidemics analysis, data visualisation and data
science, and all their dependencies.

To use the deployer, please see the guidelines below. **Do not skip any step
unless you know what you are doing.**


# About this deployer

This is the *RECON deployer* version 03.18 (indicating the release date: March
2018), which is the first stable version we distributed for deployment to the
field, or in conditions where data analysis is done without a reliable internet
connection.

We do our best to make things work - for real! But if you think you have
detected a potential bug, have requests, suggestions or questions, please post
an issue on:

[https://github.com/reconhub/deployer/issues](https://github.com/reconhub/deployer/issues)


# Using the *RECON deployer*

## Installing R / Rstudio

### Windows

Go to the `extra` folder, and install, in this order:

1. Git (`Git-2.19.0-64-bit.exe`)
2. Rtools (`Rtools34.exe`)
3. R (`R-3.5.1-win.exe`)
4. RStudio (`RStudio-1.1.456.exe`)


### Linux / MacOSX

On linux or MacOSX, it is assumed that the toolchain for package development is
already installed on the system, but we provide:

- R sources, which can be used to compile R (`R-3.5.1.tar.gz`); see guidelines
  below '*compiling R*
  
- `RStudio-1.1.456.dmg`: Rstudio for MacOSX systems, installed as usual MacOSX
  packages

- `rstudio-1.1.456-amd64.deb`: Rstudio for Linux debian-based systems (including
  Ubuntu); can by installed by typing `sudo dpkg -i rstudio-1.1.456-amd64.deb`

- `rstudio-1.1.456-x86_64.rpm`: Rstudio for Linux redhat based systems; can by
  installed by typing `sudo rpm -i rstudio-1.1.456-x86_64.rpm`

To compile and install R from the sources, open a terminal where the `.tar.gz`
file is located and type the following commands:

```{r eval = FALSE}
tar xvfz R-3.5.1.tar.gz
cd R-3.5.1
./configure
make 
sudo make install
```



<br>

## Installing packages using the *deployer*

To install packages using the *RECON deployer* instead of internet, you need
to:

1. activate the deployer by sourcing the script `activate.R`, located at the
   root of the deployer
2. install packages as usual (the deployer replaces your internet connection)


To activate the deployer on windows, the easiest approach is to open the
deployer folder, open the script `activate.R` using Rstudio, and type in the R
console:
```{r eval = FALSE}
source("activate.R")
```

you should see an output resembling:
```
 ____                            _            _ 
|  _ \ ___  ___ ___  _ __     __| | ___ _ __ | | ___  _   _  ___ _ __ 
| |_) / _ \/ __/ _ \| '_ \   / _` |/ _ \ '_ \| |/ _ \| | | |/ _ \ '__|
|  _ <  __/ (_| (_) | | | | | (_| |  __/ |_) | | (_) | |_| |  __/ |
|_| \_\___|\___\___/|_| |_|  \__,_|\___| .__/|_|\___/ \__, |\___|_|
                                       |_|            |___/

/// Examples: installing packages //

// install stable (CRAN) version of package 'outbreaks':
install.packages("outbreaks")

// install devel (github) version of package 'reportfactory':
install.packages("reportfactory")

```

You can then install CRAN packages as usual (but without internet), for
instance:

```{r eval = FALSE}
install.packages("outbreaks")
```

Non-CRAN packages are installed exactly the same way as CRAN packages; for
instance, to install `reportfactory`, use:

```{r eval = FALSE}
install.packages("reportfactory")
```

<br>
<br>
