
<div>
<img src=img/logo_deployer.png alt="recon_deployer_logo" width="600px">
</div>

<br>


# Welcome to the *RECON deployer*!

The *RECON deployer* is a project of the R Epidemics Consortium
([RECON](http://www.repidemicsconsortium.org)) aiming to provide a portable,
internet-free R environment for epidemics analysis. In a *deployer* folder, you
will find the latest version of R and associated tools (RStudio, Rtools, ...),
as well as over 1,000 packages for data analysis, including all RECON packages,
some other useful packages for epidemics analysis, data visualisation and data
science, and all their dependencies.

To use the deployer, please see the guidelines below. **Do not skip any step
unless you know what you are doing.**


# About this deployer

This is the *RECON deployer* version 03.18 (indicating the release date: March
2018), which is the first stable version we distributed for deployment to the
field, or in conditions where data analysis is done without a reliable internet
connection.

We do our best to make things work - for real! But if you think you have
detected a potential bug, have requests, suggestions or questions, please post
an issue on:

[https://github.com/reconhub/deployer/issues](https://github.com/reconhub/deployer/issues)


# Using the *RECON deployer*

## Installing R / Rstudio

### Windows

Go to the `extra` folder, and install, in this order:

1. Git (`Git-2.19.0-64-bit.exe`)
2. Rtools (`Rtools34.exe`)
3. R (`R-3.5.1-win.exe`)
4. RStudio (`RStudio-1.1.456.exe`)


### MacOS

It is assumed that command line tools and git are already installed on the
system, but we provide both R and RStudio as binaries:


- `R-3.5.1.pkg`: R installer for MacOS. Install by double-clicking and following
  the instructions.
- `RStudio-1.1.456.dmg`: Rstudio for MacOS systems, installed as usual MacOS
  apps by dragging the application to the `Applications` folder.


### Linux

It is assumed that the toolchain for package development is
already installed on the system, but we provide:

- `rstudio-1.1.456-amd64.deb`: Rstudio for Linux debian-based systems (including
  Ubuntu); can by installed by typing `sudo dpkg -i rstudio-1.1.456-amd64.deb`

- `rstudio-1.1.456-x86_64.rpm`: Rstudio for Linux redhat based systems; can by
  installed by typing `sudo rpm -i rstudio-1.1.456-x86_64.rpm`

To install R, see the next section on installing R from source:

### From Source

We have included R sources, which can be used to compile R (`R-3.5.1.tar.gz`).

To compile and install R from the sources, open a terminal where the `.tar.gz`
file is located and type the following commands:

```{r eval = FALSE}
tar xvfz R-3.5.1.tar.gz
cd R-3.5.1
./configure
make 
sudo make install
```



<br>

## Installing packages using the *deployer*

To install packages using the *RECON deployer* instead of internet, you need to:

1. Double-click on `deployer.Rproj` to open RStudio to this directory.
   (If you do not have RStudio installed, then please make sure you set your
    working directory to the deployer directory. Open R, copy and past this code,
   and select the `deployer.Rproj` file from the popup browser: 
   `setwd(dirname(file.choose()))`)
1. Activate the deployer by sourcing the script `activate.R`, located at the
   root of the deployer (use `source("activate.R")`)
2. install packages as usual (the deployer replaces your internet connection)


To activate the deployer on windows or MacOS, the easiest approach is to open
the deployer folder, double click on `deployer.Rproj` to open RStudio, and type
in the R console:

```{r eval = FALSE}
source("activate.R")
```

you should see an output resembling:

```
 ____                            _            _ 
|  _ \ ___  ___ ___  _ __     __| | ___ _ __ | | ___  _   _  ___ _ __ 
| |_) / _ \/ __/ _ \| '_ \   / _` |/ _ \ '_ \| |/ _ \| | | |/ _ \ '__|
|  _ <  __/ (_| (_) | | | | | (_| |  __/ |_) | | (_) | |_| |  __/ |
|_| \_\___|\___\___/|_| |_|  \__,_|\___| .__/|_|\___/ \__, |\___|_|
                                       |_|            |___/

/// Examples: installing packages //

// install stable (CRAN) version of package 'outbreaks':
install.packages("outbreaks")

// install devel (github) version of package 'reportfactory':
install.packages("reportfactory")

```

You can then install CRAN packages as usual (but without internet), for
instance:

```{r eval = FALSE}
install.packages("outbreaks")
```

Non-CRAN packages are installed exactly the same way as CRAN packages; for
instance, to install `reportfactory`, use:

```{r eval = FALSE}
install.packages("reportfactory")
```

## Updating a single package/fixing bugs

If you are actively maintaining a deployer in the field, it's not feasible to
create a brand-new deployerfor every small change to one package you need to
do. The script `update_deployer.R` has a function called `deploy_package()`
that will build a deployer package from source and update the repositories in
the deployer, allowing updates to take place. To use this, simply do the
following

```{r eval = FALSE}
source("update_deployer.R")
deploy_package("../incidence") # build new version of the incidence package,
                               # assuming the source code is in the directory
                               # above the deployer. 
```


<br>
<br>
